Formulating and Validating RDF Constraints Generically

### 2014-12-30 paper/RDF validation requirements/RDF validation requirements.tex
evaluation of paper 2

@@

\begin{itemize}
	\item R-70-REPEATABLE-PROPERTIES
\end{itemize}

@@

@@

\begin{itemize}
	\item R-71-CONDITIONAL-PROPERTIES
\end{itemize}

\begin{itemize}
	\item various conditions possible
  \item condition: universal quantification on object and data properties
  \item condition: existential quantification on object and data properties
  \item condition: if specific properties are present, then specific other properties also have to be present
\end{itemize}

@@

@@

\begin{itemize}
	\item R-72-RECOMMENDED-PROPERTIES
\end{itemize}

@@

\begin{itemize}
	\item get 2. list element
\end{itemize}

@@

@@

\begin{ex}
# ShEx
# ----
<IssueShape> {
    :related @<IssueShape>*
}
\end{ex}

@@

\begin{DL}
Not sure if this is expressible in DL, I think not!
\end{DL}

\an{not sure!}

@@

\begin{itemize}
  \item In most Library applications, these tend to appear in pairs, with repeatable / non-repeatable establishing max cardinality and optional / mandatory establishing min cardinality.
	\item Optional \& Non-Repeatable = [0,1]
  \item Optional \& Repeatable = [0,*]
  \item Mandatory \& Non-Repeatable = [1,1]
  \item Mandatory \& Repatable = [1,*]
\end{itemize}

\begin{DL}
Not sure if the meaning of this constraint, probably it is the same as for cardinality restrictions?
\end{DL}

\an{not sure!}

@@

@@

@@

>>>>

\newenvironment{evaluation-generic}{
  \scriptsize
  \sffamily
  \vspace{0.3cm}
	\begin{center}
  \begin{tabular}{l|c|c|c}
  \hline
  \textbf{constraint} & \textbf{property c.} & \textbf{simple c.} & \textbf{DL} \\
  \hline

\newenvironment{evaluation-generic-overview}{
  \scriptsize
  \sffamily
  \vspace{0.3cm}
	\begin{center}
  \begin{tabular}{l|c|c}
  \hline
  \textbf{Constraint Classes} & \textbf{\#} & \textbf{\%} \\
  \hline

@@

\ms{Repeatable Properties} corresponds to the requirement \ms{R-70-REPEATABLE-PROPERTIES}.

@@

The property \ms{commandsVessel} is repeatable for individuals of the class \ms{Captain}.

\begin{gcotable}
property & Captain & commandsVessel & - & $\top$ & repeatable & - \\
\end{gcotable}

@@

\ms{Conditional Properties} corresponds to the requirement \ms{R-71-CONDITIONAL-PROPERTIES}.

Multiple conditions are possible:

\begin{itemize}
  \item universal quantification on object and data properties,
  \item existential quantification on object and data properties,
  \item if specific properties are present, then specific other properties also have to be present
\end{itemize}

@@

If an individual has a \ms{parentOf} property relationship, then this individual also has a \ms{ancestorOf} property relationship.

@@

\ms{Recommended Properties} corresponds to the requirement \ms{R-72-RECOMMENDED-PROPERTIES}.

@@

get 2. list element

@@

Append a string list element 'list element' to a list the property \ms{p} of the class \ms{C} is pointing to.

\begin{gcotable}
property & C & p & - & xsd:string & append list element & 'list element' \\
\end{gcotable}

@@

\begin{ex}
# ShEx
# ----
IssueShape {
    related @IssueShape*
}
\end{ex}

\begin{DL}
$IssueShape \sqsubseteq \geq1 related.IssueShape $
\end{DL}

\begin{gcotable}
property & IssueShape & related & - & IssueShape & $\geq$ & 1 \\
\end{gcotable}

@@

If dc:date and one of its sub-properties dcterms:created or dcterms:issued are present, check that the value in dc:date is not redundant with dcterms:created or dcterms:issued for ingestion

\begin{gcotable}
property & $\top$ & dc:date & dcterms:created, dcterms:issued & - & not redundant & - \\
\end{gcotable}

@@

In most Library applications, cardinality shortcuts tend to appear in pairs, with repeatable / non-repeatable establishing maximum cardinality and optional / mandatory establishing minimum cardinality.
\begin{itemize}
	\item Optional \& Non-Repeatable = [0,1]
  \item Optional \& Repeatable = [0,*]
  \item Mandatory \& Non-Repeatable = [1,1]
  \item Mandatory \& Repatable = [1,*]
\end{itemize}
Can be expressed in DL using minimum and maximum cardinality restrictions.
We propose to use syntactic sugar instead:
\begin{gcotable}
property & C & p & - & - & optional and non-repeatable & - \\
\end{gcotable}

@@

\begin{DL}
p = COUNT (q)
\end{DL}

Context class is \ms{C}.

@@

\begin{gcotable}
property & C & p & q & - & count & - \\
\end{gcotable}

@@

\subsection{Generic Constraints}

\begin{evaluation-generic-overview}
Property Constraints & 48 & 65 (64.86) \\
Class Constraints & 17 & 23 (22.96) \\
Property and Class Constraints & 9 & 12 (12.16) \\
\hline
Simple Constraints & 46 & 62 (62.16) \\
Simple Constraints (Syntactic Sugar) & 10 & 14 (13.51) \\
Complex Constraints & 18 & 24 (24.32) \\
\hline
DL Expressible & 51 & 69 (68.92) \\
DL Not Expressible & 23 & 31 (31.08) \\
\hline
Total & 74 & 100 \\
\end{evaluation-generic-overview}

legend of the detailed evaluation:
\begin{itemize}
	\item \ms{property constraint:} property constraint ($\checkmark$) vs. class constraint (\ding{55}) vs. property and class constraints ($\sim$) 
	\item \ms{simple constraint:} simple constraint ($\checkmark$) vs. simple constraint [syntactic sugar] ($\sim$) vs. complex constraint (\ding{55})
	\item \ms{DL:} expressible in DL ($\checkmark$) vs. not expressible in DL (\ding{55})
\end{itemize}

\begin{evaluation-generic}
*Subsumption & \ding{55} & $\checkmark$ & $\checkmark$ \\ 
*Class Equivalence & \ding{55} & \ding{55} & $\checkmark$ \\ 
*Sub Properties & $\checkmark$ & $\checkmark$ & $\checkmark$ \\  
*Property Domain & $\checkmark$ & $\sim$ & $\checkmark$ \\ 
*Property Range & $\checkmark$ & $\sim$ & $\checkmark$ \\ 
*Inverse Object Properties & $\checkmark$ & $\checkmark$ & $\checkmark$ \\  
*Symmetric Object Properties & $\checkmark$ & $\checkmark$ & $\checkmark$ \\  
*Asymmetric Object Properties & $\checkmark$ & $\sim$ & $\checkmark$ \\ 
*Reflexive Object Properties & $\checkmark$ & $\sim$ & $\checkmark$ \\ 
*Irreflexive Object Properties & $\checkmark$ & $\sim$ & $\checkmark$ \\
Disjoint Properties & $\checkmark$ & $\checkmark$ & $\checkmark$ \\  
Disjoint Classes & \ding{55} & \ding{55} & $\checkmark$ \\ 
Context-Sp. Property Groups & $\sim$ & \ding{55} & $\checkmark$ \\  
Context-Sp. Inclusive OR of P. & $\sim$ & \ding{55} & $\checkmark$ \\
Context-Sp. Inclusive OR of P. Groups & $\sim$ & \ding{55} & $\checkmark$ \\ 
Recursive Queries & $\checkmark$ & $\checkmark$ & $\checkmark$ \\  
Individual Inequality & \ding{55} & \ding{55} & $\checkmark$ \\  
*Equivalent Properties & $\checkmark$ & \ding{55} & $\checkmark$ \\   
Property Assertions & $\checkmark$ & $\checkmark$ & $\checkmark$ \\  
Data Property Facets & \ding{55} & $\checkmark$ & \ding{55} \\  
Literal Pattern Matching & \ding{55} & $\checkmark$ & \ding{55} \\   
Negative Literal Pattern Matching & \ding{55} & \ding{55} & \ding{55} \\ 
*Object Property Paths & $\checkmark$ & $\checkmark$ & $\checkmark$ \\  
*Intersection & \ding{55} & $\checkmark$ & $\checkmark$ \\  
*Disjunction & \ding{55} & $\checkmark$ & $\checkmark$ \\ 
*Negation & \ding{55} & $\checkmark$ & $\checkmark$ \\ 
*Existential Quantification & $\checkmark$ & $\checkmark$ & $\checkmark$ \\  
*Universal Quantification & $\checkmark$ & $\checkmark$ & $\checkmark$ \\  
*Minimum Unqualified Cardinality & $\checkmark$ & $\checkmark$ & $\checkmark$ \\  
*Minimum Qualified Cardinality & $\checkmark$ & $\checkmark$ & $\checkmark$ \\  
*Maximum Unqualified Cardinality & $\checkmark$ & $\checkmark$ & $\checkmark$ \\  
*Maximum Qualified Cardinality & $\checkmark$ & $\checkmark$ & $\checkmark$ \\  
*Exact Unqualified Cardinality & $\checkmark$ & $\sim$ & $\checkmark$ \\ 
*Exact Qualified Cardinality & $\checkmark$ & $\sim$ & $\checkmark$ \\ 
*Transitive Object Properties & $\checkmark$ & $\checkmark$ & $\checkmark$ \\  
Context-Sp. Exclusive OR of P. & $\sim$ & \ding{55} & $\checkmark$ \\ 
Context-Sp. Exclusive OR of P. Groups & $\sim$ & \ding{55} & $\checkmark$ \\
Allowed Values & \ding{55} & $\checkmark$ & $\checkmark$ \\  
Not Allowed Values & \ding{55} & \ding{55} & $\checkmark$ \\  
Literal Ranges & \ding{55} & \ding{55} & \ding{55} \\ 
Negative Literal Ranges & \ding{55} & \ding{55} & \ding{55} \\ 
Required Properties & $\checkmark$ & $\checkmark$ & $\checkmark$ \\  
Optional Properties & $\checkmark$ & $\sim$ & $\checkmark$ \\
Repeatable Properties & $\checkmark$ & $\checkmark$ & $\checkmark$ \\  
Negative Property Constraints & $\sim$ & \ding{55} & $\checkmark$ \\ 
*Individual Equality & \ding{55} & \ding{55} & $\checkmark$ \\  
*Functional Properties & $\checkmark$ & $\checkmark$ & $\checkmark$ \\ 
*Inverse-Functional Properties & $\checkmark$ & \ding{55} & $\checkmark$ \\ 
*Value Restrictions & $\checkmark$ & $\checkmark$ & $\checkmark$ \\ 
*Self Restrictions & $\sim$ & \ding{55} & $\checkmark$ \\
Primary Key Properties & $\checkmark$ & $\sim$ & $\checkmark$ \\ 
*Class-Specific Property Range & $\checkmark$ & $\sim$ & $\checkmark$ \\ 
*Class-Sp. Reflexive Object P. & $\checkmark$ & $\checkmark$ & $\checkmark$ \\  
Membership in Controlled Vocabularies & $\sim$ & \ding{55} & $\checkmark$ \\ 
IRI Pattern Matching & \ding{55} & $\checkmark$ & \ding{55} \\   
Literal Value Comparison & $\checkmark$ & $\checkmark$ & \ding{55} \\   
Define Order & $\checkmark$ & $\checkmark$ & \ding{55} \\ 
Validation Levels & $\sim$ & $\checkmark$ & \ding{55} \\
String Operations & $\checkmark$ & $\checkmark$ & \ding{55} \\ 
\end{evaluation-generic}

\begin{evaluation-generic}
Context-Specific Valid Classes & \ding{55} & $\checkmark$ & \ding{55} \\   
Context-Specific Valid Properties & $\checkmark$ & $\checkmark$ & \ding{55} \\ 
Default Values & $\checkmark$ & $\checkmark$ & \ding{55} \\ 
Mathematical Operations & $\checkmark$ & $\checkmark$ & \ding{55} \\ 
Language Tag Matching & $\checkmark$ & $\checkmark$ & \ding{55} \\ 
Language Tag Cardinality & $\checkmark$ & $\checkmark$ & \ding{55} \\ 
Whitespace Handling & $\checkmark$ & $\checkmark$ & \ding{55} \\ 
HTML Handling & $\checkmark$ & $\checkmark$ & \ding{55} \\ 
Conditional Properties & $\checkmark$ & $\checkmark$ & $\checkmark$ \\  
Recommended Properties & $\checkmark$ & $\checkmark$ & \ding{55} \\ 
Handle RDF Collections & $\checkmark$ & $\checkmark$ & \ding{55} \\ 
Value is Valid for Datatype & $\checkmark$ & $\checkmark$ & \ding{55} \\ 
*Exploiting Class/Property Specialization Ontology Axioms & $\checkmark$ & $\checkmark$ & \ding{55} \\ 
*Cardinality Shortcuts & $\checkmark$ & $\checkmark$ & $\checkmark$ \\  
Aggregations & $\checkmark$ & $\checkmark$ & \ding{55} \\ 
\end{evaluation-generic}

%\begin{evaluation-generic}
%*Subsumption 
%*Class Equivalence
%*Sub Properties 
%*Property Domain 
%*Property Range 
%*Inverse Object Properties 
%*Symmetric Object Properties 
%*Asymmetric Object Properties 
%*Reflexive Object Properties 
%*Irreflexive Object Properties
%Disjoint Properties 
%Disjoint Classes
%Context-Sp. Property Groups 
%Context-Sp. Inclusive OR of P.
%Context-Sp. Inclusive OR of P. Groups 
%Recursive Queries 
%Individual Inequality 
%*Equivalent Properties 
%Property Assertions 
%Data Property Facets 
%Literal Pattern Matching 
%Negative Literal Pattern Matching 
%*Object Property Paths 
%*Intersection 
%*Disjunction 
%*Negation 
%*Existential Quantification 
%*Universal Quantification 
%*Minimum Unqualified Cardinality 
%*Minimum Qualified Cardinality 
%*Maximum Unqualified Cardinality 
%*Maximum Qualified Cardinality 
%*Exact Unqualified Cardinality 
%*Exact Qualified Cardinality 
%*Transitive Object Properties 
%Context-Sp. Exclusive OR of P. 
%Context-Sp. Exclusive OR of P. Groups
%Allowed Values 
%Not Allowed Values 
%Literal Ranges 
%Negative Literal Ranges 
%Required Properties 
%Optional Properties
%Repeatable Properties 
%Negative Property Constraints 
%*Individual Equality 
%*Functional Properties
%*Inverse-Functional Properties 
%*Value Restrictions
%*Self Restrictions
%Primary Key Properties 
%*Class-Specific Property Range 
%*Class-Sp. Reflexive Object P. 
%Membership in Controlled Vocabularies 
%IRI Pattern Matching 
%Literal Value Comparison 
%Define Order 
%Validation Levels 
%String Operations 
%\end{evaluation-generic}
%
%\begin{evaluation-generic}
%Context-Specific Valid Classes 
%Context-Specific Valid Properties 
%Default Values 
%Mathematical Operations 
%Language Tag Matching 
%Language Tag Cardinality 
%Whitespace Handling 
%HTML Handling 
%Conditional Properties 
%Recommended Properties 
%Handle RDF Collections 
%Value is Valid for Datatype 
%*Exploiting Class/Property Specialization Ontology Axioms 
%*Cardinality Shortcuts 
%Aggregations 
%\end{evaluation-generic}

### 2014-12-30 paper/paper 2/paper 2.tex
evaluation of paper 2

@@

We evaluated to which extend the most promising five constraint languages fulfill each requirement to formulate RDF constraints \cite{BoschNolleAcarEckert2015}.
If a constraint can be expressed in DL, we added the mapping to DL and to the generic constraint.
If a constraint cannot be expressed in DL, we added a mapping to the generic constraint.
Therefore, we show that each constraint can be mapped to a generic constraint.

\begin{evaluation}
\ms{Expr-C} (41) & 6 & 18(4) & 7(4) & 8 & 12 & 40(1)\\
\ms{Expr-R} (33) & 8 & 29 & 11 & 11(2) & 11(1) & 33 \\ 
\end{evaluation}

>>>>

\newenvironment{evaluation-generic-overview}{
  %\scriptsize
  %\sffamily
  %\vspace{0.3cm}
	\begin{center}
  \begin{tabular}{l|c|c}
  \hline
  \textbf{Constraint Classes} & \textbf{\#} & \textbf{\%} \\
  \hline

}{
  \hline
  \end{tabular}
  \linebreak
	\end{center}
}

@@

We evaluated to which extend the most promising five constraint languages fulfill each of the overall 74 requirements to formulate RDF constraints \cite{BoschNolleAcarEckert2015}.
If a constraint can be expressed in DL, we added the mapping to DL and to the generic constraint.
If a constraint cannot be expressed in DL, we added a mapping to the generic constraint.
Therefore, we show that each constraint can be mapped to a generic constraint.
The following table shows the absolute number and the relative percentage for the three dimensions to classify constraints:

\begin{evaluation-generic-overview}
Property Constraints & 48 & 65 (64.86) \\
Class Constraints & 17 & 23 (22.96) \\
Property and Class Constraints & 9 & 12 (12.16) \\
\hline
Simple Constraints & 46 & 62 (62.16) \\
Simple Constraints (Syntactic Sugar) & 10 & 14 (13.51) \\
Complex Constraints & 18 & 24 (24.32) \\
\hline
DL Expressible & 51 & 69 (68.92) \\
DL Not Expressible & 23 & 31 (31.08) \\
\hline
Total & 74 & 100 \\
\end{evaluation-generic-overview}

Constraints can be classified as \ms{property constraints} and \ms{class constraints}.
Two thirds are property constraints, one fifth class constraints, and approx. 10\% are composed of both property and class constraints.
Constraints may be either atomic (\ms{simple constraints}) or created out of simple and complex constraints (\ms{complex constraints}).
Almost two thirds are simple constraints, a quarter are complex constraints.
Almost 15 percent are complex constraints which can be formulated as simple constraints when used in terms of syntactic sugar.
Constraints can either be expressible in DL or not.
Nearly 70\% of the overall constraints are expressible in DL.
